### 2020-02-11 analysis/data_round2.Rmd

---
title: "Prag Bat Round 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)
library(corrr)
library(corrplot)
library(ggcorrplot)
library(reshape2)
library(psych)
library(readODS)
library(brms)
library(psy)

```

```{r}
## die flogenden Zeilen kannst du überspringen. Weiter gehts in Zeile 68
files <- dir("~/Work/PragBat/raw_data2/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/PragBat/raw_data2/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,34,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date,
                                               trial = as.character(trial))
  raw_data <- bind_rows(raw_data, id)
}

raw_data

```