### 2020-03-26 analysis/data_processing.Rmd
switch to Bayesian correlations

---
title: "Prag Bat"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)
library(corrr)
library(corrplot)
library(ggcorrplot)
library(reshape2)
library(psych)
library(readODS)
library(brms)
library(broom)
library(psy)
library(ggridges)

#devtools::install_github("rasmusab/bayesian_first_aid")
#library(BayesianFirstAid)
```

```{r}
#importing data
## die flogenden Zeilen kannst du überspringen. Weiter gehts in Zeile 68
# files <- dir("~/Work/PragBat/raw_data/")
# 
# raw_data <- data_frame()
# for (f in files) {
#   jf <- paste("~/Work/PragBat/raw_data/",f,sep="")
#   jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
#   date <- str_sub(jf,33,str_length(jf)-11)
#   id <- as_data_frame(jd$data$data) %>% mutate(test_date = date,
#                                                trial = as.character(trial))
#   raw_data <- bind_rows(raw_data, id)
# }
# 
# unique(raw_data$subid)
# 
# 
# task_data <- raw_data %>%
#   filter(trial != "training",
#          trial != "filler1",
#          trial != "filler2",
#          trial != "train",
#          trial != "train2")%>%
#   filter(phase == "shape" | is.na(phase))%>%
#   mutate(subid = recode(subid, `13.2` = "Vp13.2",
#                         `13.2` = "Vp13.2",
#                         `14.2` = "Vp14.2",
#                         `16.2` = "Vp16.2",
#                         `19.1` = "Vp19.1",
#                         `12.1` = "Vp12.1",
#                         `29.2` = "Vp29.2",
#                         `26.1` = "Vp26.2",
#                         `37.1` = "Vp37.1",
#                         `38.1` = "Vp38.1"
#                         ),
#          item = ifelse(task == "card_sorting", sample,leftObject))%>%
#   select(subid, subage, test_date, task,item, trial, correct)%>%
#   separate(subid, into = c("id","test_day"), sep="\\.")%>%
#   mutate(test_day = ifelse(id == "Vp12" & test_date =="2019-11-26", 2, test_day))
# 
# 
# #task_data <- read_csv("../data/data.csv")
# 
# # damit du das daten file mit dem log file (deine Liste) mergen kannst (brauchen wir für das numerische alter) legst du deine liste in den gleichen ordner in dem auch ader repository ordner liegt. Dann sollte es klappen.
# 
# log <- read_ods("../../VP Liste_Manuel.ods", sheet = 2)%>%
#   mutate(age = as.numeric(difftime(test_day_1, `birthday `, units = "days"))/365.25,
#          id = paste("Vp",`ID `, sep = ""))%>%
#   select(id, age, status)
# 
# 
# data <- task_data %>%
#   left_join(log)%>%
#   filter(status == "keep")%>%
#   mutate(check_age = ifelse(subage == substr(age,1,1),T,F),
#          subage = factor(as.numeric(substr(age,1,1))),
#          trial = as.numeric(trial))%>%
#   select(-test_date, - status, -check_age)%>%
#   mutate(trial = ifelse(test_day == "2", 
#                         ifelse(task == "card_sorting", trial + 6, trial +5),
#                         trial),
#          z_age = age - mean(age),
#          z_trial = trial - mean(trial))
# 
# write_csv(data, "../data/data.csv")

data <- read_csv("../data/data.csv")

```

# Total sample size

```{r}
data%>%
  group_by(subage)%>%
  summarise(n= length(unique(id)))%>%
  kable()

```

# Sample size with two test days

```{r}
data %>%
  group_by(subage,id)%>%
  summarise(testdays = length(unique(test_day)))%>%
  filter(testdays == 2)%>%
  summarise(complete_retest_data = length(id))%>%
  kable()
```

```{r}
# data %>%
#   group_by(id,test_day)%>%
#   summarise(n())%>%
#   arrange(id,test_day)

```

```{r}
# data%>%
#   group_by(test_day,task)%>%
#   summarise(n= length(unique(id)))
```

# Results by task binned by year

```{r}
p1 <- data%>%
  group_by(id,subage,task)%>%
  filter(task != "training")%>%
  summarise(mean = mean(correct))

p2 <- p1 %>%
  group_by(subage,task)%>%
  multi_boot_standard(col = "mean")

ggplot()+
  geom_hline(yintercept = .5, lty = 2)+
  geom_line(data = p2,aes(x = subage, y = mean, group = task), alpha = .5)+
  geom_jitter(data = p1, aes(x = subage, y = mean, col = subage), alpha = .5, width = .2, height = .01)+
  geom_pointrange(data = p2, aes(x = subage, y = mean, ymin = ci_lower, ymax = ci_upper, col = subage))+
  facet_wrap(~task)+
  labs(x = "Age Group", y = "Proportion Correct")+
  scale_color_ptol(name = "Age")+
  theme_few()+
  theme(legend.position = c(.85,.2))
```

# Models

## Card Sorting

```{r}
# card sorting

card_sorting_data <- data %>%
  filter(task == "card_sorting")

card_sorting_model1 <- brm(correct ~ z_age + z_trial + (z_trial|id) + (z_age|item),
                    data = card_sorting_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/card_sorting_model1.rds")

card_sorting_model2 <- brm(correct ~ z_trial + (z_trial|id) + (z_age|item),
                    data = card_sorting_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/card_sorting_model2.rds")

card_sorting_model1 <-readRDS("../saves/card_sorting_model1.rds")

card_sorting_model2 <-readRDS("../saves/card_sorting_model2.rds")

cs_ts <- loo_compare(card_sorting_model1, card_sorting_model2, criterion = "waic")%>%
  as.data.frame() %>%
  tibble::rownames_to_column("Model")%>%
  mutate(
    Model = recode(Model, card_sorting_model1 = "correct ~ age + trial + RE", 
                   card_sorting_model2 = "correct ~ trial + RE"),
    weight = exp(elpd_waic) / sum(exp(elpd_waic)),
    weight = round(weight, 3),
    WAIC = round(waic, 2),
    SE = round(se_waic,2))%>%
  select(Model, WAIC, SE, weight)

cs_ts%>%
  kable()

#summary(card_sorting_model1)

```

## Inforamtiveness

```{r}

inf_data <- data %>%
  filter(task == "informativeness")

inf_model1 <- brm(correct ~ z_age + z_trial + (z_trial|id) + (z_age|item),
                    data = inf_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/inf_model1.rds")

inf_model2 <- brm(correct ~ z_trial + (z_trial|id) + (z_age|item),
                    data = inf_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/inf_model2.rds")

inf_model1 <-readRDS("../saves/inf_model1.rds")

inf_model2 <-readRDS("../saves/inf_model2.rds")

inf_ts <- loo_compare(inf_model1, inf_model2, criterion = "waic")%>%
  as.data.frame() %>%
  tibble::rownames_to_column("Model")%>%
  mutate(
    Model = recode(Model, inf_model1 = "correct ~ age + trial + RE", 
                   inf_model2 = "correct ~ trial + RE"),
    weight = exp(elpd_waic) / sum(exp(elpd_waic)),
    weight = round(weight, 3),
    WAIC = round(waic, 2),
    SE = round(se_waic,2))%>%
  select(Model, WAIC, SE, weight)

inf_ts%>%
  kable()

#summary(inf_data_model1)

```

## Mutual Exclusivity

```{r}

me_data <- data %>%
  filter(task == "mutual_exclusivity")

me_model1 <- brm(correct ~ z_age + z_trial + (z_trial|id) + (z_age|item),
                    data = me_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/me_model1.rds")

me_model2 <- brm(correct ~ z_trial + (z_trial|id) + (z_age|item),
                    data = me_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/me_model2.rds")

me_model1 <-readRDS("../saves/me_model1.rds")

me_model2 <-readRDS("../saves/me_model2.rds")

me_ts <- loo_compare(me_model1, me_model2, criterion = "waic")%>%
  as.data.frame() %>%
  tibble::rownames_to_column("Model")%>%
  mutate(
    Model = recode(Model, me_model1 = "correct ~ age + trial + RE", 
                   me_model2 = "correct ~ trial + RE"),
    weight = exp(elpd_waic) / sum(exp(elpd_waic)),
    weight = round(weight, 3),
    WAIC = round(waic, 2),
    SE = round(se_waic,2))%>%
  select(Model, WAIC, SE, weight)

me_ts%>%
  kable()

#summary(me_model1)

```

## Novelty

```{r}

nov_data <- data %>%
  filter(task == "novelty")

nov_model1 <- brm(correct ~ z_age + z_trial + (z_trial|id) + (z_age|item),
                    data = nov_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/nov_model1.rds")

nov_model2 <- brm(correct ~ z_trial + (z_trial|id) + (z_age|item),
                    data = nov_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/nov_model2.rds")

nov_model1 <-readRDS("../saves/nov_model1.rds")

nov_model2 <-readRDS("../saves/nov_model2.rds")

nov_ts <- loo_compare(nov_model1, nov_model2, criterion = "waic")%>%
  as.data.frame() %>%
  tibble::rownames_to_column("Model")%>%
  mutate(
    Model = recode(Model, nov_model1 = "correct ~ age + trial + RE", 
                   nov_model2 = "correct ~ trial + RE"),
    weight = exp(elpd_waic) / sum(exp(elpd_waic)),
    weight = round(weight, 3),
    WAIC = round(waic, 2),
    SE = round(se_waic,2))%>%
  select(Model, WAIC, SE, weight)

nov_ts%>%
  kable()

#summary(nov_model1)

```

## Preference

```{r}

pref_data <- data %>%
  filter(task == "preference")

pref_model1 <- brm(correct ~ z_age + z_trial + (z_trial|id) + (z_age|item),
                    data = pref_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/pref_model1.rds")

pref_model2 <- brm(correct ~ z_trial + (z_trial|id) + (z_age|item),
                    data = pref_data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 10000)%>%
  add_criterion("waic")%>%
  saveRDS(., "../saves/pref_model2.rds")

pref_model1 <-readRDS("../saves/pref_model1.rds")

pref_model2 <-readRDS("../saves/pref_model2.rds")

pref_ts <- loo_compare(pref_model1, pref_model2, criterion = "waic")%>%
  as.data.frame() %>%
  tibble::rownames_to_column("Model")%>%
  mutate(
    Model = recode(Model, pref_model1 = "correct ~ age + trial + RE", 
                   pref_model2 = "correct ~ trial + RE"),
    weight = exp(elpd_waic) / sum(exp(elpd_waic)),
    weight = round(weight, 3),
    WAIC = round(waic, 2),
    SE = round(se_waic,2))%>%
  select(Model, WAIC, SE, weight)

pref_ts%>%
  kable()

#summary(pref_model1)

```

# Visualizing predictor for age and trial

```{r}

x <- bind_rows(
  posterior_samples(card_sorting_model1, "^b")%>%
  select(-b_Intercept)%>%
  mutate(task = "1 card_sorting"),
  posterior_samples(inf_model1, "^b")%>%
  select(-b_Intercept)%>%
  mutate(task = "4 informativeness"),
  posterior_samples(nov_model1, "^b")%>%
  select(-b_Intercept)%>%
  mutate(task = "2 novelty"),
  posterior_samples(me_model1, "^b")%>%
  select(-b_Intercept)%>%
  mutate(task = "3 mutual_exclusivity"),
  posterior_samples(pref_model1, "^b")%>%
  select(-b_Intercept)%>%
  mutate(task = "0 preference")
)%>%
  gather(estimate, value, -task)

ggplot(x, aes(y = task, x = value, fill = factor(stat(quantile))))+
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = c(0.025, 0.975)
  ) +
  facet_grid(~estimate, scales = "free_x")+
  geom_vline(xintercept = 0, col = "grey", lty = 2)+
  scale_fill_manual(values = c("white", "lightblue", "white"))+ 
  guides(fill = F)+
  theme_few()

```

# Results by task continous

```{r, cache = T}
# Files to plot data means
p_data <- data %>%
  filter(task != "training")%>%
  mutate(subage = as.numeric(as.character(subage)))%>%
  group_by(task, age,subage,id) %>%
  summarise(correct = mean(correct))

p_data2 <- p_data %>%
  group_by(task, subage) %>%
  multi_boot_standard(col = "correct")

# Files to plot individual samples from each task
samples_to_draw <- sample(1:20000, 500)

## card sorting
samples_cs <- posterior_samples(card_sorting_model1, "^b")%>%
  mutate(sample = 1:length(b_z_age))%>%
  filter(sample %in% samples_to_draw)%>%
  expand_grid(data$z_age)%>%
  mutate(y = plogis(b_Intercept + b_z_age * `data$z_age`),
         age = `data$z_age` + mean(data$age),
         task = "card_sorting")

## informativeness
samples_inf <- posterior_samples(inf_model1, "^b")%>%
  mutate(sample = 1:length(b_z_age))%>%
  filter(sample %in% samples_to_draw)%>%
  expand_grid(data$z_age)%>%
  mutate(y = plogis(b_Intercept + b_z_age * `data$z_age`),
         age = `data$z_age` + mean(data$age),
         task = "informativeness")

## mutual exclusivity
samples_me <- posterior_samples(me_model1, "^b")%>%
  mutate(sample = 1:length(b_z_age))%>%
  filter(sample %in% samples_to_draw)%>%
  expand_grid(data$z_age)%>%
  mutate(y = plogis(b_Intercept + b_z_age * `data$z_age`),
         age = `data$z_age` + mean(data$age),
         task = "mutual_exclusivity")
## novelty

samples_nov <- posterior_samples(nov_model1, "^b")%>%
  mutate(sample = 1:length(b_z_age))%>%
  filter(sample %in% samples_to_draw)%>%
  expand_grid(data$z_age)%>%
  mutate(y = plogis(b_Intercept + b_z_age * `data$z_age`),
         age = `data$z_age` + mean(data$age),
         task = "novelty")

## preference

samples_pref <- posterior_samples(pref_model1, "^b")%>%
  mutate(sample = 1:length(b_z_age))%>%
  filter(sample %in% samples_to_draw)%>%
  expand_grid(data$z_age)%>%
  mutate(y = plogis(b_Intercept + b_z_age * `data$z_age`),
         age = `data$z_age` + mean(data$age),
         task = "preference")

## combine files
samples <- bind_rows(
  samples_cs,
  samples_inf,
  samples_me,
  samples_nov,
  samples_pref
)

# File to plot means of distributions for each task

## card sorting
map_cs <- tibble(
  age = data$z_age,
  int = fixef(card_sorting_model1)[1,1],
  slope = fixef(card_sorting_model1)[2,1]
)%>%
  mutate(y = plogis(int + slope * age),
         age = age + mean(data$age),
         task = "card_sorting")

## informativeness
map_inf <- tibble(
  age = data$z_age,
  int = fixef(inf_model1)[1,1],
  slope = fixef(inf_model1)[2,1]
)%>%
  mutate(y = plogis(int + slope * age),
         age = age + mean(data$age),
         task = "informativeness")

## mutual exclusivity
map_me <- tibble(
  age = data$z_age,
  int = fixef(me_model1)[1,1],
  slope = fixef(me_model1)[2,1]
)%>%
  mutate(y = plogis(int + slope * age),
         age = age + mean(data$age),
         task = "mutual_exclusivity")

## novelty
map_nov <- tibble(
  age = data$z_age,
  int = fixef(nov_model1)[1,1],
  slope = fixef(nov_model1)[2,1]
)%>%
  mutate(y = plogis(int + slope * age),
         age = age + mean(data$age),
         task = "novelty")

## preference
map_pref <- tibble(
  age = data$z_age,
  int = fixef(pref_model1)[1,1],
  slope = fixef(pref_model1)[2,1]
)%>%
  mutate(y = plogis(int + slope * age),
         age = age + mean(data$age),
         task = "preference")

maps <- bind_rows(
  map_cs,
  map_inf,
  map_me,
  map_nov,
  map_pref
)

ggplot() +
  geom_hline(yintercept = .5, lty=2, size = 1)+
  geom_line(data = samples, aes(x = age, y = y, group = sample), size = .025, col = "grey")+
  geom_line(data = maps, aes(x = age, y = y), size = 1)+
  geom_jitter(data = p_data,aes(x = age, y= correct), width = .00, height = .01, alpha = .2)+
  geom_pointrange(data = p_data2, aes(x = subage+.5, y = mean, ymin = ci_lower, ymax = ci_upper),size = .5, position = position_dodge(width = .2), pch = 5, col = "#6866cc")+
  labs(x="Age",y="Proportion correct")+
  facet_wrap(~task, nrow = 1)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(3,5)+
  guides(alpha = F, col = F)+
  coord_fixed(ratio = 2)#+
  #ggsave("../graphs/task_results.png", width = 10, height = 2.5, scale = 1.2)
```

# Reliability

## Plot
```{r}
p3 <- data%>%
  filter(task != "training")%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  rename("Day1" = `1`,
         "Day2" = `2`)%>%
  na.omit()

ggplot(p3, aes(x = Day1, y = Day2))+
  geom_jitter(width = .05, height = .05, alpha = .5)+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.7, size = 0.5)+
  geom_smooth(method = "lm", col = "#af6300", se = T)+
  facet_grid(~task)+
  coord_fixed()+
  xlim(-.1,1.1)+
  ylim(-.1,1.1)+
  theme_few()

```

## Table
```{r}
reli_data <- data%>%
  filter(task != "training")%>%
  droplevels()%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  na.omit()%>%
  rename("Day1" = `1`,
         "Day2" = `2`)

# reli_card_sorting <- reli_data %>%
#   filter(task == "card_sorting")%>%
#   bayes.cor.test( ~Day1 + Day2, data =. )
# 
# reli_informativeness <- reli_data %>%
#   filter(task == "informativeness")%>%
#   bayes.cor.test( ~Day1 + Day2, data =. )
# 
# reli_novelty <- reli_data %>%
#   filter(task == "novelty")%>%
#   bayes.cor.test( ~Day1 + Day2, data =. )
# 
# reli_mutual_exclusivity <- reli_data %>%
#   filter(task == "mutual_exclusivity")%>%
#   bayes.cor.test( ~Day1 + Day2, data =. )
# 
# reli_preference <- reli_data %>%
#   filter(task == "preference")%>%
#   bayes.cor.test( ~Day1 + Day2, data =. )
# 
# 
# reli_table <- bind_rows(
# 
#   reli_card_sorting%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task = "card_sorting")%>%
#   select(task,mean,HDIlo,HDIup),
# 
#   reli_informativeness%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task = "informativeness")%>%
#   select(task,mean,HDIlo,HDIup),
# 
#   reli_novelty%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task = "novelty")%>%
#   select(task,mean,HDIlo,HDIup),
# 
#   reli_mutual_exclusivity%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task = "mutual_exclusivity")%>%
#   select(task,mean,HDIlo,HDIup),
# 
#   reli_preference%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task = "preference")%>%
#   select(task,mean,HDIlo,HDIup)
# )

#write_csv(reli_table, "../data/reli_table.csv")

reli_table <- read_csv("../data/reli_table.csv")

reli_table
```

```{r}
# reli_plot <- bind_rows(
#   reli_card_sorting%>%
#     as_tibble()%>%
#     mutate(task = "card_sorting"),
# 
#   reli_informativeness%>%
#     as_tibble()%>%
#     mutate(task = "informativeness"),
# 
#   reli_novelty%>%
#     as_tibble()%>%
#     mutate(task = "novelty"),
# 
#   reli_mutual_exclusivity%>%
#     as_tibble()%>%
#     mutate(task = "mutual_exclusivity"),
# 
#   reli_preference%>%
#     as_tibble()%>%
#     mutate(task = "preference")
# )
# 
# write_csv(reli_plot, "../data/reli_plot.csv")

reli_plot <- read_csv("../data/reli_plot.csv")

ggplot(reli_plot, aes(y = task, x = rho, fill = factor(stat(quantile))))+
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = c(0.025,0.5, 0.975)
  ) +
  xlim(-1,1)+
  geom_vline(xintercept = 0, col = "grey", lty = 2)+
  xlab("test - retest correlation")+
  scale_fill_manual(values = c("white", "lightblue", "lightblue", "white"))+ 
  guides(fill = F)+
  theme_few()

```

# Overview: reliability and bi-variat correlations

```{r}
cor_data <- data%>%
  filter(task != "training")%>%
  droplevels()%>%
  group_by(id,task)%>%
  summarise(mean = mean(correct))%>%
  spread(task, mean)%>%
  ungroup()%>%
  select(-id)

# cor_ci <- cor_data %>%
#   bayes.cor.test(~card_sorting + informativeness, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "card_sorting",
#          task2 = "informativeness")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_cm <- cor_data %>%
#   bayes.cor.test(~card_sorting + mutual_exclusivity, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "card_sorting",
#          task2 = "mutual_exclusivity")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_cn <- cor_data %>%
#   bayes.cor.test(~card_sorting + novelty, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "card_sorting",
#          task2 = "novelty")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_cp <- cor_data %>%
#   bayes.cor.test(~card_sorting + preference, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "card_sorting",
#          task2 = "preference")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_im <- cor_data %>%
#   bayes.cor.test(~informativeness + mutual_exclusivity, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "informativeness",
#          task2 = "mutual_exclusivity")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_in <- cor_data %>%
#   bayes.cor.test(~informativeness + novelty, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "informativeness",
#          task2 = "novelty")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_ip <- cor_data %>%
#   bayes.cor.test(~informativeness + preference, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "informativeness",
#          task2 = "preference")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_mn <- cor_data %>%
#   bayes.cor.test(~mutual_exclusivity + novelty, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "mutual_exclusivity",
#          task2 = "novelty")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_mp <- cor_data %>%
#   bayes.cor.test(~mutual_exclusivity + preference, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "mutual_exclusivity",
#          task2 = "preference")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# cor_np <- cor_data %>%
#   bayes.cor.test(~novelty + preference, data =. )%>%
#   summary()%>%
#   tidy()%>%
#   as_tibble()%>%
#   filter(.rownames == "rho")%>%
#   mutate(task1 = "novelty",
#          task2 = "preference")%>%
#   select(task1, task2 ,mean,HDIlo,HDIup)%>%
#   mutate_if(is.numeric,round, digits = 2)
# 
# 
# 
# cor <- bind_rows(
# cor_ci,
# cor_cm,
# cor_cn,
# cor_cp,
# cor_im,
# cor_in,
# cor_ip,
# cor_mn,
# cor_mp,
# cor_np
# )
# 
# write_csv(cor, "../data/bivariate_correlations.csv")

cor <- read_csv("../data/bivariate_correlations.csv")
```

```{r}
cor_plot <- bind_rows(
  cor,  
  reli_table%>%
    mutate(task1 = task,
          task2 = task)%>%
    select(task1, task2, mean, HDIlo,HDIup)%>%
    mutate_if(is.numeric,round, digits = 2)
)
```

```{r}
ggplot(cor_plot, aes(x = task2, y = task1, fill = mean))+
  geom_tile(color = "white")+
  labs(x = "", y = "")+
  scale_fill_gradient2(low = "#CC6677", high = "#117733", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation") +
   coord_fixed()+
  theme_few(base_size = 12)+
  geom_text(aes(label = mean), color = "black", size = 3) +
  theme(legend.justification = c(1, 0),
        legend.position = c(0.55, 0.7),
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

# Factor analysis

```{r}
fa_data <- p1 %>%
  spread(task, mean)%>%
  ungroup()%>%
  select(-id, -subage)%>%
  na.omit()

pca <- principal(fa_data,nfactors=2, rotate="varimax")

pca

```