### 2020-10-27 analysis/reliability.Rmd

---
title: "Prag Bat"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(brms)
```

```{r}

data <- read_csv("../data/data.csv")

```

# Total sample size

```{r}
data%>%
  group_by(subage)%>%
  summarise(n= length(unique(id)))%>%
  kable()

```

# Sample size with two test days

```{r}
data %>%
  group_by(subage,id)%>%
  summarise(testdays = length(unique(test_day)))%>%
  filter(testdays == 2)%>%
  summarise(complete_retest_data = length(id))%>%
  kable()
```

```{r}
# data %>%
#   group_by(id,test_day)%>%
#   summarise(n())%>%
#   arrange(id,test_day)

```

```{r}
# data%>%
#   group_by(test_day,task)%>%
#   summarise(n= length(unique(id)))
```

# Models

## correlations

```{r}
# card sorting

complete_data <- reli_data %>%
  group_by(subage,id)%>%
  summarise(testdays = length(unique(test_day)))%>%
  filter(testdays == 2)%>%
  pull(id)

reli_data <- data%>%
  filter(task != "training", 
         id %in% complete_data)%>%
  mutate(test_day = ifelse(test_day == 1, "day1", "day2"))

rel_model <- brm(correct ~ (test_day|task/id),
                 data = reli_data, 
                 family = bernoulli(),
                 control = list(adapt_delta = 0.9),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = F,
          iter = 2000)

summary(rel_model)

task <- as_tibble(ranef(rel_model)$task, rownames = "task")%>%
  select(task,Estimate.Intercept, Estimate.test_dayday2)%>%
  rename("Day1_int" = Estimate.Intercept,
         "Day2_int" = Estimate.test_dayday2)%>%
  mutate(task = str_replace(task, "card_sorting", "cardsorting"),
         task = str_replace(task, "mutual_exclusivity", "mutualexclusivity"))

id <- as_tibble(ranef(rel_model)$`task:id`, rownames = "id")%>%
  select(id,Estimate.Intercept, Estimate.test_dayday2)%>%
  rename("Day1" = Estimate.Intercept,
         "Day2" = Estimate.test_dayday2)%>%
  mutate(id = str_replace(id, "card_sorting", "cardsorting"),
         id = str_replace(id, "mutual_exclusivity", "mutualexclusivity"),
         Day1 = plogis(Day1),
         Day2 = plogis(Day2))%>%
  separate(id, c("task", "id"), sep = "_")

left_join(id, task)%>%
  mutate(est_1 = plogis(Day1_int + Day1),
         est_2 = plogis(Day2_int + Day2))%>%
  group_by(task)%>%
  summarise(cor = cor(est_1, est_2))

```

## reliability

```{r}
# card sorting

cor_model <- brm(correct ~ 1 + (task|id) + (task|item),
                    data = data, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = TRUE,
          iter = 2000)

summary(cor_model)

```