### 2020-10-27 analysis/summary.Rmd

---
title: "Prag Bat Summary"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(knitr)
library(ggthemes)
library(jsonlite)
library(readxl)
library(corrr)
library(corrplot)
library(ggcorrplot)
library(reshape2)
#library(psych)
library(readODS)
library(brms)
library(psy)
library(tidyboot)
library(lubridate)
library(ggpubr)

```

```{r, include=FALSE}
data <- bind_rows(
  read_csv("../data/data_r1.csv")%>%filter(task != "training")%>%mutate(round = "R1"),
  read_csv("../data/data_r2.csv")%>%mutate(round = "R2")
  )%>%
  mutate(round = factor(round),
         id = paste(round, id, sep = "_"))

```

```{r, include=FALSE}
data%>%
  group_by(subage)%>%
  summarise(n= length(unique(id)))
```
# Objective

Assess reliability of and relations between different pragmatic inference tasks which we have developed over the years.

# Sample size

Our goal was to test at least 20 children per age group and round twice. Children were tested in a kindergarten in Leipzig.

```{r}
data %>%
  group_by(subage,id,round)%>%
  summarise(testdays = length(unique(test_day)))%>%
  group_by(round,subage)%>%
  summarise(n = length(unique(id)),
            complete_retest_data = sum(testdays == 2))%>%
  kable()
```

# Results by Task

All tasks had 5 trials, except for card sorting, which had 6. All subjects got the tasks in the same order and the same version of each task.

Round 1 included the following tasks:

* Informativeness (same setup and stimuli as in MCC)
* Preference (as in MCC)
* Novelty (as in SPIN)
* Mutual exclusivity (as in SPIN, with fewer items)
* Card sorting (DCCS sensu Zelazo, 2006)

Round 1 included the following tasks:

* Simple informativeness (same setup and stimuli as in Frank & Goodman, 2014)
* Ad-hoc implicature (same stimuli as Yoon & Frank, 2019)
* Discourse continuity (as in DISCON)
* Mutual exclusivity (as in SPIN, with fewer items, same as R1)
* Card sorting (DCCS sensu Zelazo, 2006, same as R1)

For mutual exclusivity and card sorting we have a direct replication. The plot below shows all data, including data from children who were tested only once.

```{r}
p1 <- data%>%
  group_by(round, id,age, subage,task)%>%
  filter(task != "training")%>%
  summarise(mean = mean(correct))

p2 <- p1 %>%
  group_by(round,subage,task)%>%
  tidyboot_mean(column = mean)%>%
  mutate(chance = ifelse(task == "discourse_continuity", 1/3, 1/2))

ggplot()+
  geom_hline(data = p2, aes(yintercept = chance), lty = 2)+
  geom_smooth(data = p1, aes(x = age, y = mean, lty = round), method = "lm", col = "black", size = 1)+
  geom_jitter(data = p1, aes(x = age, y = mean, pch = round), alpha = .5, width = .05, height = .01)+
  geom_pointrange(data = p2, aes(x = as.numeric(as.character(subage))+.5, y = mean, ymin = ci_lower, ymax = ci_upper, col = factor(subage), pch = round), position = position_dodge(width = .5))+
  facet_wrap(~task)+
  labs(x = "Age Group", y = "Proportion Correct")+
  scale_color_ptol(name = "Age")+
  ylim(-0.05, 1.05)+
  theme_few()+
  theme(legend.position = c(.85,.1), legend.direction = "horizontal")

```

# Reliability

Based on simple Pearson correlations on the data aggregated by subject, task and test day.

```{r}
reli <- data%>%
  filter(task != "training")%>%
  droplevels()%>%
  group_by(round,id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  na.omit()%>%
  rename("Day1" = `1`,
         "Day2" = `2`)%>%
  group_by(round,task)%>%
  summarize(reli = cor.test(Day1,Day2)$estimate,
            lci = cor.test(Day1,Day2)$conf.int[1],
            uci = cor.test(Day1,Day2)$conf.int[2],
            p = cor.test(Day1,Day2)$p.value)%>%
  mutate_if(is.numeric, round, digits = 2)

reli%>%
  kable()
```

# Correlations between tasks

Based on simple Pearson correlations on the data aggregated by subject and task.

```{r, fig.align="center"}
cor_r1 <- data%>%
  filter(round == "R1")%>%
  droplevels()%>%
  group_by(id,task)%>%
  summarise(mean = mean(correct))%>%
  spread(task, mean)%>%
  ungroup()%>%
  select(-id)%>%
  corrr::correlate(diagonal = reli%>%filter(round == "R1")%>%pull(reli))%>%
  gather(task, cor, -rowname)%>%
  mutate(cor = replace(cor, duplicated(cor), NA))%>%
  mutate_if(is.numeric, round, digits = 2)%>%
  na.omit()%>%
  mutate(round = "R1")

cor_r2 <- data%>%
  filter(round == "R2")%>%
  droplevels()%>%
  group_by(id,task)%>%
  summarise(mean = mean(correct))%>%
  spread(task, mean)%>%
  ungroup()%>%
  select(-id)%>%
  corrr::correlate(diagonal = reli%>%filter(round == "R2")%>%pull(reli))%>%
  gather(task, cor, -rowname)%>%
  mutate(cor = replace(cor, duplicated(cor), NA))%>%
  mutate_if(is.numeric, round, digits = 2)%>%
  na.omit()%>%
  mutate(round = "R2")

cor <- bind_rows(
  cor_r1,
  cor_r2
)

ggarrange(
ggplot(cor_r1, aes(x = rowname, y = task, fill = cor))+
  geom_tile(color = "white")+
  labs(x = "", y = "")+
  scale_fill_gradient2(low = "#CC6677", high = "#117733", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation") +
   coord_fixed()+
  theme_few()+
  geom_text(aes(label = cor), color = "black", size = 3) +
  ggtitle("Round 1")+
  theme(
        legend.position = "right",
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)),

ggplot(cor_r2, aes(x = rowname, y = task, fill = cor))+
  geom_tile(color = "white")+
  labs(x = "", y = "")+
  scale_fill_gradient2(low = "#CC6677", high = "#117733", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation") +
   coord_fixed()+
  theme_few()+
  geom_text(aes(label = cor), color = "black", size = 3) +
  ggtitle("Round 2")+
  theme(
        legend.position = "right",
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)),
common.legend = T, legend = "top"

)
```