### 2020-11-25 analysis/data_round2.Rmd
add file for reli sensu rouder and haaf

---
title: "Prag Bat Round 2"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(knitr)
library(ggthemes)
library(jsonlite)
library(readxl)
library(corrr)
library(corrplot)
library(ggcorrplot)
library(reshape2)
#library(psych)
library(readODS)
library(brms)
library(psy)
library(tidyboot)
library(lubridate)

```

```{r}
files <- dir("~/Work/Local/PragBat/raw_data2/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/Local/PragBat/raw_data2/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,34,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date,
                                               trial = as.character(trial))
  raw_data <- bind_rows(raw_data, id)
}

task_data <- raw_data%>%
  filter(trial != "training",
         trial != "filler1",
         trial != "filler2",
         trial != "train",
         trial != "train1",
         trial != "train2",
         task != "training")%>%
  filter(phase == "shape" | is.na(phase))%>%
  separate(subid, into = c("id","test_day"), sep="_")

log <- read_xlsx("../../../../Cloud/PragBat/vpl_pragBat2.xlsx", sheet = 1)%>%
  mutate(age = as.numeric(difftime(as.Date(as.character(test_day_1),format="%Y%m%d"), as.Date(as.character(birthday),format="%Y%m%d"), units = "days"))/365.25,
         age_num_month = lubridate::time_length(difftime(as.Date(as.character(test_day_1),format="%Y%m%d"), as.Date(as.character(birthday),format="%Y%m%d")), "month"),
          age_month = round(age_num_month)/12,
         id = paste("VP",ID, sep = ""))%>%
  select(id, age, age_month, status)

data <- task_data %>%
  left_join(log)%>%
  filter(status == "keep")%>%
  mutate(check_age = ifelse(subage == substr(age,1,1),T,F),
         subage = factor(as.numeric(substr(age,1,1))),
         trial = as.numeric(trial))%>%
  mutate(trial = ifelse(test_day == "2", 
                        ifelse(task == "card_sorting", trial + 6, trial +5),
                        trial),
         z_age = age - mean(age),
         z_trial = trial - mean(trial),
         item = ifelse(task == "card_sorting", sample, ifelse(task == "discourse_continuity", target1, leftObject)))%>%
    select(id, test_day, subage,age,age_month, task,item, trial, correct, z_age, z_trial)

 # write_csv(data, "../data/data_r2.csv")
```

```{r}
data%>%
  group_by(subage)%>%
  summarise(n= length(unique(id)))
```

```{r}
unique(data$id)

```

```{r}
data %>%
  group_by(subage,id)%>%
  summarise(testdays = length(unique(test_day)))%>%
  filter(testdays == 2)%>%
  summarise(complete_retest_data = length(id))

data %>%
  group_by(subage)%>%
  filter(test_day == 2)%>%
  summarise(complete_retest_data = length(unique(id)))
```

```{r}
p1 <- data%>%
  group_by(id,age, subage,task)%>%
  filter(task != "training")%>%
  summarise(mean = mean(correct))

p2 <- p1 %>%
  group_by(subage,task)%>%
  tidyboot_mean(column = mean)%>%
  mutate(chance = ifelse(task == "discourse_continuity", 1/3, 1/2))

ggplot()+
  geom_hline(data = p2, aes(yintercept = chance), lty = 2)+
  geom_smooth(data = p1, aes(x = age, y = mean), method = "lm", col = "black", size = 1)+
  geom_jitter(data = p1, aes(x = age, y = mean), alpha = .5, width = .05, height = .01)+
  geom_pointrange(data = p2, aes(x = as.numeric(as.character(subage))+.5, y = mean, ymin = ci_lower, ymax = ci_upper, col = subage))+
  facet_wrap(~task)+
  labs(x = "Age Group", y = "Proportion Correct")+
  scale_color_ptol(name = "Age")+
  ylim(-0.05, 1.05)+
  theme_few()+
  theme(legend.position = c(.85,.2))

```

```{r}
p3 <- data%>%
  filter(task != "training")%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  rename("Day1" = `1`,
         "Day2" = `2`)%>%
  na.omit()

ggplot(p3, aes(x = Day1, y = Day2))+
  geom_jitter(width = .05, height = .05, alpha = .5)+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 0.7, size = 0.5)+
  geom_smooth(method = "lm", col = "#af6300", se = T, size = 0.5)+
  facet_grid(~task)+
  coord_fixed()+
  xlim(-.1,1.1)+
  ylim(-.1,1.1)+
  theme_few()

```

```{r}
reli <- data%>%
  filter(task != "training")%>%
  droplevels()%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  na.omit()%>%
  rename("Day1" = `1`,
         "Day2" = `2`)%>%
  group_by(task)%>%
  summarize(reli = cor.test(Day1,Day2)$estimate,
            lci = cor.test(Day1,Day2)$conf.int[1],
            uci = cor.test(Day1,Day2)$conf.int[2],
            p = cor.test(Day1,Day2)$p.value)%>%
  mutate_if(is.numeric, round, digits = 2)
```

```{r}
cor <- data%>%
  filter(task != "training")%>%
  droplevels()%>%
  group_by(id,task)%>%
  summarise(mean = mean(correct))%>%
  spread(task, mean)%>%
  ungroup()%>%
  select(-id)%>%
  corrr::correlate(diagonal = reli$reli)%>%
  gather(task, cor, -rowname)%>%
  mutate(cor = replace(cor, duplicated(cor), NA))%>%
  mutate_if(is.numeric, round, digits = 2)%>%
  na.omit()

ggplot(cor, aes(x = rowname, y = task, fill = cor))+
  geom_tile(color = "white")+
  labs(x = "", y = "")+
  scale_fill_gradient2(low = "#CC6677", high = "#117733", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation") +
   coord_fixed()+
  theme_few(base_size = 12)+
  geom_text(aes(label = cor), color = "black", size = 3) +
  theme(legend.justification = c(1, 0),
        legend.position = c(0.55, 0.7),
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

# Discourse continuity by category

```{r}

con_data <- task_data%>%
  filter(task == "discourse_continuity")%>%
  filter(!is.na(correct))

pc1 <- con_data%>%
  group_by(target1,subage, id) %>%
  summarise(correct = mean(correct))

pc2 <- pc1 %>%
  group_by(target1,subage) %>%
  tidyboot_mean(column = correct)

ggplot()+
  geom_hline(yintercept = 1/3, lty=2)+
  geom_jitter( data = pc1, aes(x = subage, y = correct, col = target1), position = position_jitterdodge(jitter.width = .1,jitter.height = .025, dodge.width = .75 ), alpha = .3)+
  geom_pointrange(data = pc2, aes(x = subage, y = mean, ymin = ci_lower, ymax = ci_upper, col = target1), position = position_dodge(width = .75), pch = 4)+
  #facet_grid(~experiment)+
  labs(x="Age",y="Proportion correct")+
  theme_few()+
  scale_color_colorblind(name = "Category")
```

# Data file Spin-off

```{r}
item_info <- data_frame(
  task = c(rep("mutual_exclusivity",10),rep("simple_inf",10),rep("ad_hoc_implicature",10),rep("discourse_continuity",10),rep("card_sorting",12)),
  trial = c(rep(1:10,4),1:12), 
  item2 = c(rep(c("carrot","garlic","eggplant","duck","horseshoe"),2),rep(c("bear","robot","rocket","truck","dino"),2), rep(c("box","chair","table","plate","house"),2),rep(NA,22))
)

retest_ids <- data %>%
  filter(test_day == 2)%>%
  group_by(id)%>%
  summarise(n())%>%
  pull(id)

data %>%
  filter(id %in% retest_ids)%>%
  select(-z_age, -z_trial)%>%
  left_join(item_info)%>%
  mutate(item = ifelse(task == "mutual_exclusivity" | task == "simple_inf" | task == "ad_hoc_implicature", item2,item))%>%
  select(id, test_day, subage, age, age_month, task, trial, item, correct )%>%
  mutate(c_age_month = age_month - min(age_month),
         c_age = age - min(age))%>%
  write_csv("../../../Spin-off/spin-off/data/data.csv")
```