### 2020-11-25 analysis/reliability.Rmd

---
title: "Prag Bat"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(brms)
library(tidybayes)
library(ggpubr)
```

```{r}

data <- bind_rows(
data_r1 <- read_csv("../data/data_r1.csv")%>%mutate(id = paste(id,"r1",sep = "_")) ,
data_r2 <- read_csv("../data/data_r2.csv")%>%mutate(id = paste(id,"r2",sep = "_"))
)%>%
  filter(task !="training")

```

# Total sample size

# Sample size with two test days

```{r}
data %>%
  group_by(task,id)%>%
  summarise(testdays = length(unique(test_day)))%>%
  filter(testdays == 2)%>%
  summarise(complete_retest_data = length(id))
```

# Simple correaltions

```{r}
reli <- data%>%
  filter(task != "training")%>%
  droplevels()%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  na.omit()%>%
  rename("Day1" = `1`,
         "Day2" = `2`)%>%
  group_by(task)%>%
  summarize(reli = cor.test(Day1,Day2)$estimate,
            lci = cor.test(Day1,Day2)$conf.int[1],
            uci = cor.test(Day1,Day2)$conf.int[2],
            p = cor.test(Day1,Day2)$p.value)%>%
  mutate_if(is.numeric, round, digits = 2)

reli
```

```{r}
p_rel_data <- data%>%
  filter(task != "training")%>%
  droplevels()%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  na.omit()%>%
  rename("day1" = `1`,
         "day2" = `2`)

ggplot(p_rel_data, aes(x = day1, y = day2))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(alpha = .5)+
  facet_wrap( ~ task)+
  stat_cor(method = "pearson",  aes(x = day1, y = day2), size = 2)+
  theme_minimal()+
  theme(aspect.ratio = 1)
```

# Models

## Combined model

```{r}
complete_data <- data %>%
  group_by(subage,id)%>%
  summarise(testdays = length(unique(test_day)))%>%
  filter(testdays == 2)%>%
  pull(id)

reli_data <- data%>%
  filter(task != "training", 
         id %in% complete_data)%>%
  mutate(test_day = ifelse(test_day == 1, "day1", "day2"),
         z_trial = scale(trial))

rel_model <- brm(correct ~ (1|test_day) + (1| task) + (1|id),
                 data = reli_data, 
                 family = bernoulli(),
                 control = list(adapt_delta = 0.9),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = F,
          iter = 5000)

summary(rel_model)

draws1 <- rel_model %>%
  spread_draws(b_Intercept, day[test_day,], task[`test_day:test_day`,],id[`test_day:test_day:id`,])
#   separate(id, c("id","round"), sep = "_", extra = "merge")
# 
# draws2 <- rel_model %>%
#   spread_draws(b_Intercept,`r_id:test_day`[`id:test_day`,])%>%
#   separate(`id:test_day`, c("id","go"), sep = "_", extra = "merge")%>%
#   separate(go, c("round","day"), sep = "_", extra = "merge")
# 
# draws3 <- rel_model %>%
#   spread_draws(b_Intercept,`r_id:test_day:task`[`id:test_day:task`,])%>%
#   separate(`id:test_day:task`, c("id","go"), sep = "_", extra = "merge")%>%
#   separate(go, c("round","go"), sep = "_", extra = "merge")%>%
#   separate(go, c("day","task"), sep = "_", extra = "merge")
# 
# draws <- draws1 %>%
#   left_join(draws2) %>%
#   left_join(draws3)

p_re_est <- draws1%>%
  mutate(perf = b_Intercept + r_id + r_test_day + r_task)%>%
  group_by(id, test_day,task )%>%
  median_qi(perf)%>%
  mutate(perf = plogis(perf))

saveRDS(p_re_est,"../saves/p_re_est.rds")

p_data <- data %>%
  filter(id %in% complete_data)%>%
  group_by(task, test_day,id)%>%
  summarise(mean = mean(correct))%>%
  mutate(day = paste("day",test_day,sep = ""))

ggplot()+
  geom_point(data = p_re_est, aes (x = id, y = perf), alpha = .5)+
  geom_point(data = p_data, aes (x = id, y = mean), alpha = .5, col = "firebrick")+
  facet_grid(day~task)+
  theme_minimal()

p_cor <- p_re_est %>%
  select(task, id, test_day,perf)%>%
  spread(test_day, perf)%>%
  group_by(task)%>%
  summarise(cor = cor(day1, day2))%>%
  left_join(reli)

p_re_est %>%
  select(task, id, test_day,perf)%>%
  spread(test_day, perf)%>%
  ggplot(aes(x = day1, y = day2))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(alpha = .5)+
  geom_point(data = p_rel_data, aes(x = day1, y = day2), alpha = .5, col = "firebrick")+
  facet_wrap( ~ task)+
  stat_cor(method = "pearson",  aes(x = day1, y = day2), size = 2)+
  theme_minimal()+
  theme(aspect.ratio = 1)

```
# Ignore below 
## Separate models

```{r}
reli_data_1 <- reli_data%>%
  filter(test_day == "day1")

rel_model_1 <- brm(correct ~ (z_trial|id/task),
                 data = reli_data_1, 
                 family = bernoulli(),
                 control = list(adapt_delta = 0.9),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = F,
          iter = 5000)

summary(rel_model_1)

draws_rel1_1 <- rel_model_1 %>%
  spread_draws(b_Intercept,r_id[id,])%>%
  separate(id, c("id","round"), sep = "_", extra = "merge")

draws_rel1_2 <- rel_model_1 %>%
  spread_draws(b_Intercept,`r_id:task`[`id:task`,])%>%
  separate(`id:task`, c("id","go"), sep = "_", extra = "merge")%>%
  separate(go, c("round","task"), sep = "_", extra = "merge")

draws_rel1 <- draws_rel1_1 %>%
  left_join(draws_rel1_2)%>%
  mutate(perf = b_Intercept + r_id + `r_id:task`)%>%
  group_by(id,task )%>%
  median_qi(plogis(perf))%>%
  mutate(day1 = `plogis(perf)`)

```

```{r}
reli_data_2 <- reli_data%>%
  filter(test_day == "day2")

rel_model_2 <- brm(correct ~ (z_trial|id/task),
                 data = reli_data_2, 
                 family = bernoulli(),
                 control = list(adapt_delta = 0.9),
          sample_prior = F,
          chains = 4,
          cores = 4,
          save_all_pars = F,
          iter = 5000)

summary(rel_model_2)

draws_rel2_1 <- rel_model_2 %>%
  spread_draws(b_Intercept,r_id[id,])%>%
  separate(id, c("id","round"), sep = "_", extra = "merge")

draws_rel2_2 <- rel_model_2 %>%
  spread_draws(b_Intercept,`r_id:task`[`id:task`,])%>%
  separate(`id:task`, c("id","go"), sep = "_", extra = "merge")%>%
  separate(go, c("round","task"), sep = "_", extra = "merge")

draws_rel2 <- draws_rel2_1 %>%
  left_join(draws_rel2_2)%>%
  mutate(perf = b_Intercept + r_id + `r_id:task`)%>%
  group_by(id,task )%>%
  median_qi(plogis(perf))%>%
  mutate(day2 = `plogis(perf)`)

```

```{r}
p_cor %>%left_join(
  draws_rel1 %>%
  select(id, task, day1)%>%
  left_join(draws_rel2%>%
  select(id, task, day2))%>%
    group_by(task)%>%
    summarise(cor_2_models = cor(day1, day2))
    )

```

```{r}
draws_rel1 %>%
  select(id, task, day1)%>%
  left_join(draws_rel2%>%
  select(id, task, day2))%>%
  ggplot(aes(x = day1, y = day2))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_point(alpha = .5)+
  geom_point(data = p_rel_data, aes(x = day1, y = day2), alpha = .5, col = "firebrick")+
  facet_wrap( ~ task)+
  stat_cor(method = "pearson",  aes(x = day1, y = day2), size = 2)+
  theme_minimal()+
  theme(aspect.ratio = 1)
```