### 2020-11-25 analysis/reliability_rouder_haaf.Rmd

---
title: "Prag Bat reliability"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(brms)
library(tidybayes)
library(ggpubr)
```

# Loading data
```{r}

raw_data <- bind_rows(
read_csv("../data/data_r1.csv")%>%mutate(id = paste(id,"r1",sep = "_")) ,
read_csv("../data/data_r2.csv")%>%mutate(id = paste(id,"r2",sep = "_"))
)%>%
  filter(task !="training")

complete_ids <- raw_data %>%
  group_by(subage,id)%>%
  summarise(testdays = length(unique(test_day)))%>%
  filter(testdays == 2)%>%
  pull(id)

data <- raw_data%>%
  filter(task != "training", 
         id %in% complete_ids)%>%
  mutate(test_day = ifelse(test_day == 1, "day1", "day2"),
         z_trial = scale(trial))

```
# Descriütives

```{r}
data %>%
  group_by(task)%>%
  summarise(n = length(unique(id)))
```

# Simple correaltions

```{r}
data%>%
  droplevels()%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  group_by(task)%>%
  summarize(reli = cor.test(day1,day2)$estimate,
            lci = cor.test(day1,day2)$conf.int[1],
            uci = cor.test(day1,day2)$conf.int[2],
            p = cor.test(day1,day2)$p.value)%>%
  mutate_if(is.numeric, round, digits = 2)

```
## Visualize simple correlations
```{r}
data%>%
  droplevels()%>%
  group_by(id,task, test_day)%>%
  summarise(mean = mean(correct))%>%
  spread(test_day, mean)%>%
  ggplot( aes(x = day1, y = day2))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_count(alpha = .5)+
  facet_wrap( ~ task)+
  stat_cor(method = "pearson",  aes(x = day1, y = day2), size = 2)+
  theme_minimal()+
  theme(aspect.ratio = 1)
```

# Model by task sensu Rouder and Haaf

## Task: Mutual exclusivity
```{r}
me_data <- data %>%
  filter(task == "mutual_exclusivity")
```