### 2022-02-19 paper/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{An individual differences perspective on the development of pragmatic reasoning}
\author{Manuel Bohn\textsuperscript{1}, Michael Henry Tessler\textsuperscript{2,3}, Clara Kordt\textsuperscript{4}, Tom Hausmann\textsuperscript{5}, \& Michael C. Frank\textsuperscript{6}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={An individual differences perspective on the development of pragmatic reasoning},
  pdfauthor={Manuel Bohn1, Michael Henry Tessler2,3, Clara Kordt4, Tom Hausmann5, \& Michael C. Frank6},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Individual differences in pragmatic reasoning}
\keywords{keywords\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

We are very thankful to Stella Christie for sharing the material for the relational match-to-sample task with us.

M. Bohn received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Sklodowska-Curie grant agreement no. 749229. M. H. Tessler was funded by the National Science Foundation SBE Postdoctoral Research Fellowship Grant No.~1911790. M. C. Frank was supported by a Jacobs Foundation Advanced Research Fellowship and the Zhou Fund for Language and Cognition. The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

The authors made the following contributions. Manuel Bohn: Conceptualization, Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael Henry Tessler: Analysis, Writing - Review \& Editing; Clara Kordt: Data collection; Tom Hausmann: Data collection; Michael C. Frank: Conceptualization, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Manuel Bohn, Max Planck Institute for Evolutionary Anthropology, Deutscher Platz 6, 04103 Leipzig, Germany. E-mail: \href{mailto:manuel_bohn@eva.mpg.de}{\nolinkurl{manuel\_bohn@eva.mpg.de}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany\\\textsuperscript{2} DeepMind, London, UK\\\textsuperscript{3} Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology\\\textsuperscript{4} Martin Luther University Halle-Wittenberg\\\textsuperscript{5} Brandenburg Medical School Theodor Fontane\\\textsuperscript{6} Department of Psychology, Stanford University}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{facets-of-pragmatic-reasoning}{%
\section{Facets of pragmatic reasoning}\label{facets-of-pragmatic-reasoning}}

\hypertarget{the-challenges-of-studying-individual-differences}{%
\section{The challenges of studying individual differences}\label{the-challenges-of-studying-individual-differences}}

\hypertarget{the-current-study}{%
\section{The current study}\label{the-current-study}}

\hypertarget{study-1}{%
\section{Study 1}\label{study-1}}

Methods and sample size were pre-registered at \url{https://osf.io/6a723}. All analysis scripts and data files can be found in the following repository: \url{https://github.com/manuelbohn/pragBat}. The same repository also contains the code to run the experiments.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

For Study 1, we collected data from 48 children (\(m_{age}\) = 3.99, range\(_{age}\): 3.10 - 4.99, 23 girls) of which 41 were tested twice. For most children, the two test sessions were two days apart; the longest time difference was six days. Children came from an ethnically homogeneous, mid-size German city (\textasciitilde550,000 inhabitants, median income €1,974 per month as of 2020); were mostly monolingual and had mixed socioeconomic backgrounds. The study was approved by an internal ethics committee at the Max Planck Institute for Evolutionary Anthropology. Data was collected between November 2019 and January 2020.

\hypertarget{material-and-methods}{%
\subsection{Material and Methods}\label{material-and-methods}}

The study was presented as an interactive picture book on a tablet computer. The tasks were programmed in \texttt{HTML/JavaScript} and run in a web browser. Pre-recorded sound files were used to address the child (one native German speaker per animal). Children responded by touching objects on the screen. Children were tested in a quiet room in their daycare or in a separate room in a child laboratory. An experimenter guided the child through the study, selecting the different tasks and advancing within each task. In the beginning of the study, children completed a touch training to familiarize themselves with selecting objects. After a short introduction to the different animal characters, children completed the following six tasks. Figure \ref{fig:fig1} shows screenshots for each task and the order in which they were presented.

\hypertarget{training}{%
\subsubsection{Training}\label{training}}

An animal was standing on a pile between two tables. On each table, a familiar object was located. The animal asked the child to give them one of the objects (e.g., ``Can you give me the car''). the objects were chosen so that children of the youngest age group would easily understand them (car and ball). This procedure familiarized the child with the general logic of the animals making requests and the child touching objects. There were two training trials.

\hypertarget{mutual-exclusivity}{%
\subsubsection{Mutual exclusivity}\label{mutual-exclusivity}}

This task was directly taken from Bohn, Tessler, Merrick, and Frank (2021). The task layout and the procedure was the same as in the training. In each trial, one object was a novel object (drawn for the purpose of this study) while the other one was likely to be familiar to children. Both object types changed from trial to trial. Following Bohn et al. (2021), the familiar objects varied in terms of the likelihood that they would be familiar to children in the age range (carrot, duck, eggplant, garlic, horseshoe). For example, we assumed that most 3-year-olds would recognize a carrot, whereas fewer children would recognize a horseshoe. The animal always used a novel non-word (e.g., gepsa) in their request. We reasoned that children would identify the novel object as the referent of the novel word because they assumed the animal would have used the familiar word if they wanted to request the familiar object. Children's response was thus coded as correct if they selected the novel object. There were five trials, with the side on which the novel object appeared pseudo-randomized.

\hypertarget{informativeness-inference}{%
\subsubsection{Informativeness inference}\label{informativeness-inference}}

The task was directly taken from Bohn, Tessler, Merrick, and Frank (2022). Th animal was standing between two trees with objects hanging in them. In one tree, there were two objects (type A and B) and in the other tree there was only one (type B). The animal turned to the tree with the two objects and labelled one of the objects. It was unclear from the animal's utterance, which of the two objects they were referring to. We assumed that children would map the novel word onto the object of type A because they expected the animal to turn to the tree with only the object of type B if their intention was to provide a label for an object of type B. Next, the trees were replaced by new ones, one of which carried an object of type A and the other of type B. The animal then said that one of the trees had the same object as they labelled previously (using the same label) and asked the child to touch the tree. We coded as correct if the child selected the tree with the object of type A. The first two trials were training trials, in which there was only one object in each tree. There were five test trials. The location of the tree with the two objects in the beginning of each trial was pseudo-randomized and so was the location of the objects when the new trees appeared.

\hypertarget{speaker-preference}{%
\subsubsection{Speaker preference}\label{speaker-preference}}

This task was also taken from Bohn et al. (2022). The animal was standing between the two tables, each of which had a novel object (drawn for the purpose of the study) on it. The animal turned to one table, pointed at the object and said that they very much liked this object (using a pronoun instead of a label). Next, the animal turned to the other table and said that they really did not like the object (again, using a pronoun and no label). Then the animal turned towards the participant and used a novel label to request an object in an excited tone. We assumed that children would track the animal's preference and identify the previously liked object as the referent. Thus, we coded as correct if the child selected the object the animal expressed preference for. There were five test trials. The location of the preferred object as well as whether the animal first expressed liking or disliking was pseudo-randomized across trials

\hypertarget{discourse-novelty}{%
\subsubsection{Discourse novelty}\label{discourse-novelty}}

This task was taken from Bohn et al. (2021). Once again, the animal was standing between the two tables. One table was empty whereas there was a novel object on the other table. The animal turned towards the empty table and commented on its emptiness. Next, the animal turned to the other table and commented (in a neutral tone) on the presence of the object (not using a label). The animal then briefly disappeared. In the absence of the animal a second novel object appeared on the previously empty table. Then the animal returned and, facing the participant, asked for an object in an excited tone. We assumed that children would track which object was new to the ongoing interaction and identify the object that was new in context as the referent. We coded as correct when children selected the object that appeared later. There were five test trials. The location of the empty table and whether the animal first commented on the presence or absence of an object was pseudo-randomized across trials

\hypertarget{card-sorting}{%
\subsubsection{Card sorting}\label{card-sorting}}

This task was modeled after Zelazo (2006). The child saw to cards, a blue rabbit on the left and a red boat on the right. The experimenter introduced the child to the color game they would be playing next. In this game, all blue cards (irrespective of object depicted) would go to the left card and all red cards to the right. Next, a third card appeared in the middle of the screen (red rabbit or blue boat) and the experimenter demonstrated the color sorting by moving the card to the one with the same color. After a second demonstration trial, the child started to do the color sorting by themselves. After six trials, the experimenter said that they were now going to play a different game, the shape game, according to which all rabbits would go to the card with the rabbit (left) and all boats to the card with the boat (right). The experimenter repeated these instructions once and without any demonstration the child continued with the sorting according to the new rule. There were six test trials. The shape on the card was pseudo-randomized across trials. We only coded the trials after the rule change and coded as correct when the child sorted according to shape.

Each child received exactly the same version of each task and completed the tasks in the same order, with the same order on the two days. This ensured comparability of performance across children.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/figure1}

}

\caption{Overview of the tasks used in Study 1 to 3. Pictures show screenshots from each task. The vertical order corresponds to the order of presentation in each study. The colors group the tasks along the (Assumed) cognitive processes involved. Blue: utterance-based inferences. Red common ground/discourse-based inferences. Green: Executive functions. Green: Analogical reasoning.}\label{fig:fig1}
\end{figure}

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

We analysed the data in three steps. First we investigated developmental effects in each task, then we assessed re-test reliability and finally, we analysed relations between the tasks.

All analysis were run in \texttt{R} (R Core Team, 2018) version 4.1.2. Regression models were fit as Bayesian generalized linear mixed models (GLMM) using the function \texttt{brm} from the package \texttt{brms} (Bürkner, 2017). We used default priors for all analysis.

To estimate developmental effects in each task, we fit a GLMM predicting correct responses (0/1) by age (centered at the mean) and trial number (also centered). The model included random intercepts for each participant and random slopes for trial within participants (model notation in \texttt{R}: \texttt{correct\ \textasciitilde{}\ age\ +\ trial\ +\ (trial\textbar{}id)} ). For each task, we inspected and visualized the posterior distribution (mean and 95\% Credible Interval (CrI)) for the age estimate.

We assessed re-test reliability in two ways. First, for each task we computed the proportion of correct trials for each individual in the two test sessions and then used Pearson correlations to quantify re-test reliability. Second, we used a GLMM based approach suggested by Rouder and Haaf (2019). Here, a GLMM was fitted to the trial-by-trial data for each task with a fixed effect of age, a random intercept for each participant and a random slope for test day (\texttt{correct\ \textasciitilde{}\ age\ +\ (0+test\_day\textbar{}id)})\footnote{The notation \texttt{0+test\_day} yields a separate intercept estimate for each test day and subject instead of an intercept estimate for day 1 and a slope for the difference between day 1 and day 2. As a consequence, the model estimates the correlation between the two test days instead of a correlation between an intercept and the slope for test day.}. The model yields a participant specific estimate for each test day and also estimates the correlation between the two. This correlation can be interpreted as the re-test reliability. This approach has several advantages. First, it uses the trial-by-trial data and avoids information loss that comes with data aggregation. Second, it uses hierarchical shrinkage to obtain better participant specific estimates. Finally, it allows us to get an age-independent estimate for reliability. One worry when assessing re-test reliability in developmental studies is that re-test correlations can be high because of domain general cognitive gains and not because of task-specific individual differences. By including age as a fixed effect in the model, the estimates for each participant are independent of age and so is the correlation between estimates for the two test days -- the re-test reliability.

Finally, we used that aggregated data from both test days for each participant and task to compute Pearson correlations between the different tasks. Given the small sample size in Study 1, this part of the analysis is mostly exploratory.

\hypertarget{results}{%
\subsection{Results}\label{results}}

We found developmental effects in most of the tasks. Figure \ref{fig:fig2} shows the data and visualizes the developmental trajectories based on the model. Figure \ref{fig:fig3} shows the model estimates for age. In the mutual exclusivity task, performance was reliably above chance level and increased with age. For informative inference, the pattern was quite different: Performance was at chance level with only minor developmental gains. In the speaker preference task, performance was again clearly above chance with developmental gains resulting in a ceiling effect for older children. In the discourse novelty task, performance was also above chance with no clear developmental effects. The card sorting task showed the strongest developmental effects with younger children performing largely below chance and older children performing above chance.

Re-test reliability was high for most tasks (see Figure \ref{fig:fig2}). Raw correlations between the two test sessions was above .7 for mutual exclusivity, speaker preference and discourse novelty. With .62 it was slightly lower for card sorting. The model based -- age independent -- reliability estimates yielded similar results suggesting that the tasks did capture task specific individual differences. A notable exception was the informativeness inference task, which was not reliable according to any of the methods of computing re-test correlations. We suspect the overall low variation in performance to be responsible for this.

Most correlations between the tasks were low and ranged between \emph{r} = -0.2 and 0.2 (see Figure \ref{fig:fig2}). A notable exception was the correlation between mutual exclusivity and card sorting with \emph{r} = 0.31 (95\% CI{[}0.03 - 0.55{]}.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/figure2}

}

\caption{Results by task for studies 1 to 3. Each panel shows the results for one task. Regression lines show the predicted developmental trajectories (with 95\% CrI) based on by-task GLMMs, with the line type indicating the study. Colored points show age group means (with 95\% CI based on non-parametric bootstrap) with the different shapes corresponding to the different studies. Light shapes show the mean performance for each subject by study. Dotted line shows level of performance expected by chance.}\label{fig:fig2}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{./figures/figure3}

}

\caption{Model estimates (with 95\% CrI) for age based on GLMMs for each task and study.}\label{fig:fig3}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/figure4}

}

\caption{xxx.}\label{fig:fig4}
\end{figure}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

Study 1 showed that the different tasks were age appropriate and reliable. A notable exception was the informativeness inference task which generated to systematic variation in the age range we studied here. Correlations between the tasks were generally low, with the notable exception of the relation between mutual exclusivity and card sorting. Given the small sample size, we want to avoid overly strong claims, however, it was interesting to see that the relation between the two tasks tapping into common ground based inferences (speaker preference and discourse novelty) were -- if anything -- negatively correlated.

\hypertarget{study-2}{%
\section{Study 2}\label{study-2}}

Based on these results od Study 1, we compiled a new set of tasks for Study 2. We retained the mutual exclusivity and card sorting tasks because of the interesting relation between the two found in Study 2. We simplified the informativeness inference task to be more age appropriate with the hope to induce more variation in performance. We removed the speaker preference and discourse novelty tasks -- despite their excellent re-test reliability -- because they seemed to be unrelated to one another and also unrelated to the other tasks. The new tasks focused on ad-hoc implicature and discourse continuity. As noted in the introduction, we had theoretical reasons to expected the ad-hoc implicature task to be related to the mutual exclusivity and informativeness inference tasks. We had no such strong predictions for the discourse continuity task.

Methods and sample size were pre-registered at \url{https://osf.io/hp9f7}. Data, analysis scripts and experiment code can be found in the associated online repository.

\hypertarget{participants-1}{%
\subsection{Participants}\label{participants-1}}

Participants for Study 2 were recruited from the same general population. We collected data from 54 children (\(m_{age}\) = 3.97, range\(_{age}\): 3.09 - 4.93, 24 girls) of which 40 were tested twice. The two test sessions were again two days apart; the longest time difference was 14 days. Data was collected between March and October 2020.

\hypertarget{material-and-methods-1}{%
\subsection{Material and Methods}\label{material-and-methods-1}}

The general setup and mode of presentation was the same as in Study 1. We added two new tasks and modified the informativeness inferences task, which we will describe in detail below. The mutual exclusivity and card sorting tasks were the same as in Study 1.

\hypertarget{informativeness-inference-1}{%
\subsubsection{Informativeness inference}\label{informativeness-inference-1}}

The general structure of the task was the same as in Study 1, however, we replaced the stimuli by the ones used by Frank and Goodman (2014). We suspected that children did not treat the novel objects hanging in the tree as properties of the tree but the tree as a ``container'' for the novel objects. The alleged inference, however, relies on seeing the objects as properties of the referent. With the new stimuli, we emphasized that the novel objects were mere properties of the referent by making the referent more salient and more different across trials. The animal was located between two identical objects, which had different properties (see Figure \ref{fig:fig1}). For example, the child saw two bears, one with a Pharaoh-style crown and a match in its hand, the other only with the match. The animal then turned to the object with the two properties and described it by referring to one of the properties (e.g., a bear with a {[}non-word{]}). Next, the objects disappeared, and the same objects re-appeared but this time, each of them had only one property (e.g., one bear with a crown, the other with the match). The animal then asked which of these objects had the aforementioned property (e.g., which bear has a {[}non-word{]}). We coded as correct if the child selected the object with the property that was unique to the object during labeling. The first two trials were training trials, in which each object only had one property. There were five test trials. The location of the object with the two properties in the beginning of each trial was pseudo-randomized and so was the location of the properties when the new objects appeared.

\hypertarget{discourse-continuity}{%
\subsubsection{Discourse continuity}\label{discourse-continuity}}

This task was directly taken from Bohn, Le, Peloquin, Köymen, and Frank (2021). Children were told that they were going to visit the animals at their place. The animal greeted the child and told them that they would show them their things. During exposure trials, the child saw three objects from three different categories (e.g., train (vehicle), drum (instrument), orange (fruit); see Figure \ref{fig:fig1}). The animal named of the objects and asked the child to touch it. On the next exposure trial, the child saw three new objects but from the same categories (e.g., bus (vehicle), flute (instrument), apple (fruit)). The animal asked the child to touch the object from the same category as previously (only naming the object, not the category). There were 5 such exposure trials. On the following test trial, the animal used a pronoun to refer to one of the objects (i.e., can you touch \emph{it}). We assumed that children would use the exposure trials to infer that the animal was talking about a certain category and would use this knowledge to identify the referent of the pronoun.Children received five test trials, each with a different category as the target. The position of the objects in exposure trials as well as test trials was pseudo-randomized.

\hypertarget{ad-hoc-implicature}{%
\subsubsection{Ad-hoc implicature}\label{ad-hoc-implicature}}

This task used the general procedure and stimuli developed in Yoon and Frank (2019). The animal was located in a window, looking out over two objects (see Figure \ref{fig:fig1}). Both objects were of the same kind, but had different properties. As properties we chose objects that were well known to children of that age range. One object had one property (A) and while the other had two (A and B). For example, objects were lunchboxes, one with an orange and the other with an orange and an apple. The animal then asked the child to hand them their object which was the one with the property that both objects shared (A). We assumed that children would pick the object with only property A because they expected the animal to name property B if they had wanted to refer to the object with both properties. There were five test trials, preceded by two training trials in which the objects did not share a common property. The positioning of the objects (left and right) was pseudo-randomized

\hypertarget{analysis-1}{%
\subsection{Analysis}\label{analysis-1}}

We used the same methods to analyse the data as in Study 1.

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

We found substantial developmental gains in all five tasks (Figure \ref{fig:fig2} and \ref{fig:fig3}). For mutual exclusivity and ad-hoc implicature performance was above chance across the entire age range. For the informativeness inference and discourse continuity tasks, performance was close to chance for younger children and reliably above it for older children. Like in Study 1, we found the strongest developmental effect for card sorting, with performance below chance for 3-year-olds and above chance for 4-year-olds.

Re-test reliability based on aggregated data was good for all tasks with most estimates around 0.7. The model-based reliability estimates were similar, with lower values for ad-hoc implicature and higher ones for discourse continuity. Notably, the informativeness inference task showed a much-improved re-test reliability compared to Study 1.

Correlations between tasks were generally higher compared to Study 1. In fact, confidence intervals for correlation coefficients were not overlapping with 0 except for the correlation between the discourse continuity and informativeness inference tasks (Figure \ref{fig:fig4}. Once again, we found the strongest relation between card sorting and mutual exclusivity (\emph{r} = 0.60, 95\% CI{[}0.40 - 0.75{]}). Other notable relations were those between card sorting and informativeness inference (\emph{r} = 0.47, 95\% CI{[}0.23 - 0.65{]}) as well as between ad-hoc implicature and informativeness inference (\emph{r} = 0.49, 95\% CI{[}0.25 - 0.67{]}).

\hypertarget{discussion-1}{%
\subsection{Discussion}\label{discussion-1}}

In Study 2 we found good results from a measurement perspective: all tasks had acceptable re-test reliability. This included the informativeness inference task which had deficits in that respect in Study 1. Higher average performance and increased variability suggest that our changes to the stimuli did make the task easier for children.

Like in Study 1, we found a relatively strong correlation between the mutual exclusivity and card sorting tasks. This corroborates the idea that these tasks share common processes. We also found substantial relations between the three utterance-based inference tasks (mutual exclusivity, ad-hoc implicature, informativeness inference). Furthermore, the correlations between tasks were lower when discourse continuity was involved -- though the numerical difference was minimal.

\hypertarget{study-3}{%
\section{Study 3}\label{study-3}}

In Study 3, we focused explicitly on the relations between the different tasks. In particular, we explored the idea that the three utterance-based inference tasks share common cognitive processes. Once again, we also included the card sorting task and added a new task of analogical reasoning for which we did not expect strong relations with the other tasks. To be able to test these predictions, we collected data from a comparatively larger sample of children.

The reliability estimates from Study 1 and 2 helped us plan the sample size for Study 3. The focal tasks had a re-test reliability around 0.7. Because the highest plausible correlation between two tasks is the product of their reliabilities (higher correlations would mean that the task is more strongly related to a different task than to itself), the highest we could expect were correlations between two tasks around 0.7 * 0.7 = 0.49. We planned our sample so that we could detect correlations between two tasks of 0.3 with 95\% power\footnote{The first author drafted a pre-registration and shared it with the last author but forgot to register it at OSF. Thus, the study was not officially pre-registered.}. Data, analysis scripts and experiment code can be found in the associated online repository.

\hypertarget{participants-2}{%
\subsection{Participants}\label{participants-2}}

For Study 3, we collected data from 126 children (\(m_{age}\) = 4.00, range\(_{age}\): 3.00 - 5.02, 74 girls) from the same general population. Data was collected between June and November 2021.

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

From Study 2, we used the the mutual exclusivity, ad-hoc implicature, informativeness inference and card\_sorting tasks. We added the relational match-to-sample task, which we now describe in more detail.

\hypertarget{realtional-match-tosample}{%
\subsubsection{Realtional match-to--sample}\label{realtional-match-tosample}}

The task was modeled after and used the original stimuli from Christie and Gentner (2014). The child saw three cards, one on top (the sample) and two at the bottom (the potential matches; see Figure \ref{fig:fig1}). The experimenter guided the child through the study and read out the instructions. The child was instructed to match the sample card to one of the lower ones based on similarity, that is, they were instructed to pick the card that was ``like'' the sample. All cards had two geometical shapes on them.

\hypertarget{analysis-2}{%
\subsection{Analysis}\label{analysis-2}}

Conf FA - evaluate: PPP, RMSAE. model comparison to alternative models based on WAIC (the lower the better) - out of sample predictive accuracy

\hypertarget{results-2}{%
\subsection{Results}\label{results-2}}

\hypertarget{discussion-2}{%
\subsection{Discussion}\label{discussion-2}}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{./figures/figure5}

}

\caption{xxx.}\label{fig:fig5}
\end{figure}

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-bohn_le_peloquin_koymen_frank_2020}{}}%
Bohn, M., Le, K. N., Peloquin, B., Köymen, B., \& Frank, M. C. (2021). Children's interpretation of ambiguous pronouns based on prior discourse. \emph{Developmental Science}, \emph{24}(3), e13049.

\leavevmode\vadjust pre{\hypertarget{ref-bohn2021young}{}}%
Bohn, M., Tessler, M. H., Merrick, M., \& Frank, M. C. (2021). How young children integrate information sources to infer the meaning of words. \emph{Nature Human Behaviour}, \emph{5}(8), 1046--1054.

\leavevmode\vadjust pre{\hypertarget{ref-bohn_tessler_merrick_frank_2019}{}}%
Bohn, M., Tessler, M. H., Merrick, M., \& Frank, M. C. (2022). Predicting pragmatic cue integration in adults' and children's inferences about novel word meanings. \emph{Journal of Experimental Psychology: General}.

\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.

\leavevmode\vadjust pre{\hypertarget{ref-christie2014language}{}}%
Christie, S., \& Gentner, D. (2014). Language helps children succeed on a classic analogy task. \emph{Cognitive Science}, \emph{38}(2), 383--397.

\leavevmode\vadjust pre{\hypertarget{ref-frank2014inferring}{}}%
Frank, M. C., \& Goodman, N. D. (2014). Inferring word meanings by assuming that speakers are informative. \emph{Cognitive Psychology}, \emph{75}, 80--96.

\leavevmode\vadjust pre{\hypertarget{ref-R-base}{}}%
R Core Team. (2018). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-rouder2019psychometrics}{}}%
Rouder, J. N., \& Haaf, J. M. (2019). A psychometrics of individual differences in experimental tasks. \emph{Psychonomic Bulletin \& Review}, \emph{26}(2), 452--467.

\leavevmode\vadjust pre{\hypertarget{ref-yoon2019role}{}}%
Yoon, E. J., \& Frank, M. C. (2019). The role of salience in young children's processing of ad hoc implicatures. \emph{Journal of Experimental Child Psychology}, \emph{186}, 99--116.

\leavevmode\vadjust pre{\hypertarget{ref-zelazo2006dimensional}{}}%
Zelazo, P. D. (2006). The dimensional change card sort (DCCS): A method of assessing executive function in children. \emph{Nature Protocols}, \emph{1}(1), 297--301.

\end{CSLReferences}

\endgroup

\end{document}